The Habitable Zone Inner Edge Discontinuity: Probing the Imprint of Runaway Greenhouse Climates on Exoplanet Demographics

### 2023-04-19 src/tex/ms.tex
revise abstract

% Abstract
\begin{abstract}
Key atmospheric volatiles are distributed between terrestrial planetary interiors and atmospheres during magma ocean epochs in the course of planetary accretion.
Long-term magma ocean regimes on rocky exoplanets inside the orbital runaway greenhouse threshold -- the inner edge of the classical habitable zone -- may thus offer insights into the physical and chemical processes that distinguish potentially habitable worlds from others.
Thermal stratification of super-runaway planets is expected to inflate the transit radii relative to sub-runaway planets, potentially providing observational access to the runaway greenhouse transition in the form of a ``habitable zone inner edge discontinuity'' in radius/density space.
Here, we assess the ability of ground- and space-based telescopes to test this hypothesis using \bioverse, a simulation framework that leverages contextual information from the overall planet population.
We find that the demographic imprint of super-runaway climates should be detectable with high-precision transit photometry for planet populations $\gtrsim$ 100 planets if at least 10\% of these planets undergo a super-runaway climate.
We argue that in the near future, ESA's \plato\ mission will be the most promising survey to detect and constrain habitable zone inner edge discontinuity.
We identify the key mission design drivers that enable a statistically sound detection, and discuss survey strategies that optimize the statistical power of combined datasets.
Observational detection of the runaway climate transition will provide crucial insights into the distribution of atmospheric volatiles among low-mass exoplanets, and inform us about which stars harbor the nearest potentially habitable worlds.
\end{abstract}

>>>>

% Abstract
\begin{abstract}
Long-term magma ocean phases on rocky exoplanets orbiting within the runaway greenhouse regime -- the inner edge of the classical habitable zone -- may offer insights into the physical and chemical processes that distinguish potentially habitable worlds from others.
Thermal stratification of super-runaway planets is expected to significantly inflate their transit radii, potentially providing observational access to the runaway greenhouse transition in the form of a ``habitable zone inner edge discontinuity'' in radius/density space.
Here, we assess the ability of ground- and space-based telescopes to test this hypothesis using \bioverse, a simulation framework that leverages contextual information from the overall planet population.

We find that the demographic imprint from super-runaway climates is likely detectable with high-precision transit photometry for planet samples of size $\gtrsim$~100 if at least \SI{\sim 10}{\percent} of planets inside of the habitable zone inner edge harbor super-runaway climates.
Our survey simulations suggest that in the near future, ESA's \plato\ mission will be the most promising survey to probe the habitable zone inner edge discontinuity.
We identify the key mission design drivers that enable a statistically sound detection and discuss survey strategies that optimize the statistical power of combined datasets.
Observational constraints on the runaway climate transition will provide crucial insights into the distribution of atmospheric volatiles among low-mass exoplanets and inform us about which stars harbor the nearest potentially habitable worlds.
\end{abstract}
